\plaintitle

### 2018-07-18 paper_corrected.tex
Corrected paper and made latexdiff
corrected most of the mistakes
need some revision
added latexdiff

% Copyright
% \CopyrightYear{2016}
% %\setcopyright{acmcopyright}
% \setcopyright{acmlicensed}
% %\setcopyright{rightsretained}
% %\setcopyright{usgov}
% %\setcopyright{usgovmixed}
% %\setcopyright{cagov}
% %\setcopyright{cagovmixed}
% % DOI
% \doi{http://dx.doi.org/10.475/123_4}
% % ISBN
% \isbn{123-4567-24-567/08/06}
% %Conference
% \conferenceinfo{CHI'16,}{May 07--12, 2016, San Jose, CA, USA}
% %Price
% \acmPrice{\$15.00}

@@

\begin{abstract}
  Smart Textiles and body worn controllers (Wearables) have been analyzed, mostly individually, using different technologies and studying different aspects. Although there were some tests concerning everyday life, these controllers are not used very often for several reasons. This paper aims to show some of them and analyze, instead, the opportunities that  may lead to better development and improve the feeling that a user has when interacting in an uncommon way with a device. One of the main threats to the adoption of Wearables that this paper will discuss will be the reasons for social unacceptability. Other topics will be the methods to review social acceptability and possible solutions to the problems concerning the usage of Wearables in public, taking into account the technology-related tradeoffs.
  To illustrate this issue, different Wearables are going to be presented - specifically concentrating on the social aspects of the devices. Finally, this paper discusses further development of social acceptability of Wearables in the future.
\end{abstract}
% \category{H.5.m.}{Information Interfaces and Presentation
%   (e.g. HCI)}{Miscellaneous} \category{See
%   \url{http://acm.org/about/class/1998/} for the full list of ACM
%   classifiers. This section is required.}{}{}
%
% \keywords{\plainkeywords}
%
% \section{Introduction} \label{test:intro}
%
% Here there's the intro.
% \\ \textbf{Bold text} is done with
% \begin{verbatim}
%   \textbf{Bold text}
% \end{verbatim}
% \emph{emphatized text} is done with
% \begin{verbatim}
%   \emph{emphatized text}
% \end{verbatim}
% New line is made with double backslash
% \begin{verbatim}
%   line \\ another line
% \end{verbatim}
% \subsection{A subsection}
% Text of the subsection is shown here.
% \subsubsection{This is a subsubsection}
% We should not go further in sub sectioning, so we shall not use \textit{paragraph} and \textit{subparagraph}
% % References must be the same font size as other body text.
\section{Introduction}
%
% quick overview over the topic. Philipp already mentioned that his abstract might be a good starting point.
% Also we should give an introduction on the respective parts of the paper and give reason for the research we did (why is social acceptability an issue…)

Touch gestures are a form of interaction that is used by almost everyone daily in our society. Whether Smartphone, Tablet, or Smartwatch - many mobile devices are nowadays controlled by this input technique to master different tasks. High demands are directly related to the fast and intuitive use of these devices. The combination of different electronic devices can help achieve these requirements to a higher extent. For example, smart watches are already used to control mobile phones or various home electronics (e.g. controlling lamps). The development of innovative input devices, which can be integrated into clothing or even applied to the user's skin, offers the potential to communicate with electronic devices in completely new ways and make interaction even more intuitive. In this paper, we especially focus on touch input devices that are attached to the user's skin or are integrated into clothing. This can be used to control different electronic devices.
Because these technologies create new and unusual forms of interaction, we will also look at third-party perceptions of a user’s interactions with a wearable E-textile interface. Social conventions play an important role in the development of these new forms of interaction. The willingness to make a gesture depends largely on how appropriate these interactions look and feel in public. \cite{touch-wrist} The usability of these technologies must also be examined: to support the user as efficiently, effectively and satisfactorily as possible, the designers must consider the interaction with the system in addition to the placement of these body-worn controllers.

\section{Showcase of technology examples}
%
% Maybe first a definition of body-worn controllers and a not, that we want to concentrate on Touch-Input.
% I think it is a good idea to pick a few examples which were mentioned in the papers and discuss them briefly.
% For the first part (just the showcase) it is sufficient to just write down the basic technology (with Source): So what it is? How does is work (very briefly)? What it is used for?
% We can get more into detail in the pros/cons section below
The body worn controllers\footnote{input sources made by joining small electric devices with everyday worn clothing} %footnote
allow to have bigger controlling surfaces ready to be used without needing to take out the main device. Moreover, these controllers allow users to express input gestures, which are comparable to everyday gestures. This allows the user to express emotions while using the device. If on a touchscreen, the feedback of the input is based only on vibration and visual input. With body-worn controllers, the input surface is placed over the skin, which allows direct feedback to the user, including the emotion put into performing a gesture. An example made by Wiegel et Al. %in More Than Touch
\cite[p.185]{more-touch} is “Uncomfortable interactions were chosen for actions that are very important and not reversible”, which shows how the gesture chosen to perform an action were related to the consequence that this action may lead to.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{jogwheel.png}
  \caption{\emph{Jogwheel} as presented in \protect\cite{touch-wrist}}
  \label{fig:1}
\end{figure}
%
%
% here we can add everything to e-Textiles
Another type of E-textile is the one presented by SmartSleeve \cite{smart-sleeve}, where the fabric is an input surface made of alternating conductive and insulating textile bands. Then another pressure-sensitive fabric is set under the first one, beneath which there is another fabric alternating conductive and non-conductive bands. This layer is positioned orthogonally to the top layer. This allows the fabric to register both the position and the intensity of a touch with a matrix representation. The sensor density is 1.66 sensors per square inch, which allows for a moderate sensing capability. However, since the system is working with the short-circuit principle, and the wires are not insulated, skin moisture can lead to false position sensing. To avoid this, the authors have put SmartSleeve over a long-sleeved running shirt. This solution, however, becomes unsuitable in summer for obvious reasons. %maybe better written
%
Project Jaquard \cite{jacquard}, instead, utilizes a capacitive system where the fabric is one layer textile, in which some of the wires in the weft and warp are electrical wires. As the touch sensing is made with capacitance measure, this system is not vulnerable to errors due to the skin moisture. Different materials have been investigated to create the electrical wires, both for simplicity of the creation of the fabric and the necessity of flexibility needed for comfort. The material chosen was copper because it has similar properties to cotton, silk and wool when the wire diameter is 50µm.
%
% \subsection{SimpleSkin: Towards Multipurpose Smart Garments}
The Simple Skin \cite{simpleskin} is an example of a modular smart textile garment, created for commercially viable multi-purpose garments. It combines multiple sensing modalities, namely capacitive sensing for head and wrist movement , resistive sensing for arm movement\cite{vogel4} and bio-impedance sensing for heart rate. This allows optional smartphone interaction. This placement is intended to detect the angles of the joints to analyse posture, monitor vital signs, and offer input on the socially accepted region of the lower arm with the wrist sensor. Finally, the smartphone grants the options of a regular input device or a supportive device to supply the user with  information. This modular design allows for the separation of various components such as sensing, acquisition hardware, processing, and software. This allows the use of various combinations of different technologies on the same garment, and enables modifications of the underlying fabric with detachable hardware. \cite{simpleskin}

@@

The perception of technology use in public spaces can be different in every culture, because the standards of social behaviour vary in many countries. This also should be considered in the development of new interactive systems, which is why it also needs to be part of the research of social acceptance. Campbell et al. have shown in their study on the public use of mobile phones, that there are indeed differences in this field. For example, Campbell et al. discovered that in Japan the use of mobile phones in buses or sidewalks is less accepted by society than in a restaurant. In many other cultures, this is rather the opposite. \cite{mobile-phones} These results lead to the conclusion that there are also cultural differences in perception when interacting with body-worn controllers which need to be explored.

@@

\subsection{State of social acceptability of wearables}
``...as basic needs are met, attention shifts to higher-order needs. In the case of wearable technologies, the field has matured to the point where basic needs like physical comfort, accessibility, and usability are met, higher-order needs for social acceptance and self-actualization become increasingly imperative.’’ \cite[p. 4162]{social-comfort} %(\cite{social-comfort} page 4162)
This quote underlines the current challenge Wearables have to face to succeed in the long term. Since as mentioned by Dunne, Lucy E., et al. the basic needs are figured out for Wearables they now have to please higher needs - such as social acceptability. To illuminate the current state some results of current researches are presented here.

@@

Result “Don’t Mind me Touching my Wrist” \cite{touch-wrist}: The "Jogwheel" \cite{touch-wrist} mentioned above was used in an online survey to evaluate the social acceptance of an E-textile interface by showing a demo video to the participants. The results showed gender and cultural differences in this field. There were large gender-specific differences, especially in body placement. For example, while the placement of the interface on a man's chest looked very accessible and natural to the participants, the same part of a woman's body was perceived as “disgusting”. \cite{touch-wrist} The participants from the USA found the interaction with the interface at the waist by a man more embarrassing than by a woman. The answers of the South Koreans were exactly reversed here, which indicates to cultural differences. \cite{touch-wrist}

@@

\subsection{Evaluation of placement areas for body-worn controllers }
The single most prominent factor mentioned in the respective research projects, besides the type of gesture, was the body region where the Wearable was placed. The figure above shows the preferred regions in the individual studies. When comparing the researches there are two main observations to be made. First there is a strict aversion against socially suggestive regions for Wearables. These regions include genitalia, the bottom and, especially when asking female subjects, the upper torso area, including the sternum and collarbone.
Second the all around preferred area for Wearable seems to be the forearm and wrist region - especially of the non dominant hand. \cite{touch-wrist,pinstripe,social-comfort,belt} %[1,6,7,8]
This may be due to the fact that interaction with technologies on these body regions is already established and accepted in society through other devices such as watches and fitness bracelets. However, these body regions are also easy to reach and offer sufficient space for various gestures, which makes these body regions very suitable for body-worn controllers in the future as well.
\section{Outlook on social acceptability of wearables in the future}
When looking into the future of Wearables one of the most interesting parts is the social acceptability. Since physical comfort, accessibility, and usability are for the most part figured out \cite{social-comfort}.
One, if not the most, important part of social acceptability are the performed gestures. Currently there is always a tradeoff which needs to be balanced when designing gestures, which are meant to be used in public situations. On one hand natural gestures are inherently more unobtrusive and socially acceptable. On the other hand these gestures are far more difficult to track and evaluate precisely, since they are easily confusable with everyday gestures that were not meant to be an input \cite{social-comfort}.
One approach to solve this problem is improving on better technology and tracking to avoid both confusing inputs and social discomfort at the same time. Another approach could be focussed around the social vocabulary more than around the gestures themselves.
``Because there does not currently exist a standard `vocabulary’ of gesture, it is correspondingly more difficult for viewers to match a perceived gesture with a previously-understood meaning.'' \cite[p. 4160]{social-comfort} %page 4160
Looking at this quote the other solution seems obvious. With the further adaptation of a novel technology there comes a better understanding of the gestures and finally a new social vocabulary regarding it. There were similar developments when the bluetooth headset \cite{social-comfort} or the Sony Walkman \cite{touch-wrist} was established.
In conclusion, the social acceptance of body-worn controllers can be increased through further developments and by bringing them more and more in contact with the public.

>>>>

%Copyright
\CopyrightYear{2018}
% \setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}
% % DOI
% \doi{http://dx.doi.org/10.475/123_4}
% % ISBN
% \isbn{123-4567-24-567/08/06}
% %Conference
% \conferenceinfo{CHI'16,}{May 07--12, 2016, San Jose, CA, USA}
% %Price
% \acmPrice{\$15.00}

@@

\begin{abstract}
  Smart Textiles and body worn controllers (Wearables) have been analyzed, mostly individually, using different technologies and studying different aspects. Although there were some tests concerning everyday life, these controllers are not used very often for several reasons. This paper aims to show some of these reasons and analyze, instead, the opportunities that  may lead to better development and improve the feeling that a user has when interacting with a device in a way that's not common along the rest of the population. One of the main threats to the adoption of Wearables that this paper will discuss will be the reasons for social unacceptability. Other topics will be the methods to review social acceptability and possible solutions to the problems concerning the usage of Wearables in public, taking into account the technology-related tradeoffs.
  To illustrate this issue, different Wearables are going to be presented - specifically concentrating on the social aspects of the devices. Finally, this paper discusses further development of social acceptability of Wearables in the future.
\end{abstract}
% \category{H.5.m.}{Information Interfaces and Presentation
%   (e.g. HCI)}{Miscellaneous} \category{See
%   \url{http://acm.org/about/class/1998/} for the full list of ACM
%   classifiers. This section is required.}{}{}
%
% \keywords{\plainkeywords}
%
% \section{Introduction} \label{test:intro}
%
% Here there's the intro.
% \\ \textbf{Bold text} is done with
% \begin{verbatim}
%   \textbf{Bold text}
% \end{verbatim}
% \emph{emphatized text} is done with
% \begin{verbatim}
%   \emph{emphatized text}
% \end{verbatim}
% New line is made with double backslash
% \begin{verbatim}
%   line \\ another line
% \end{verbatim}
% \subsection{A subsection}
% Text of the subsection is shown here.
% \subsubsection{This is a subsubsection}
% We should not go further in sub sectioning, so we shall not use \textit{paragraph} and \textit{subparagraph}
% % References must be the same font size as other body text.
\section{Introduction}
%
% quick overview over the topic. Philipp already mentioned that his abstract might be a good starting point.
% Also we should give an introduction on the respective parts of the paper and give reason for the research we did (why is social acceptability an issue…)

Touch gestures are a form of interaction that is used a lot in our society.\footnote{To date, just counting only Android and iOS devices, more than 3 billion devices are active monthly, https://www.apple.com/newsroom/2018/02/apple-reports-first-quarter-results/ and https://twitter.com/Google/status/864890655906070529} Whether Smartphone, Tablet, or Smartwatch - many mobile devices are nowadays controlled by this input technique to master different tasks. High demands are directly related to the fast and intuitive use of these devices. The combination of different electronic devices can help achieve these requirements to a higher extent. For example, smart watches are already used to control mobile phones or various home electronics (e.g. controlling lamps). The development of innovative input devices, which can be integrated into clothing or even applied to the user's skin, offers the potential to communicate with electronic devices in completely new ways and make interaction even more intuitive. In this paper, we especially focus on touch input devices that are attached to the user's skin or are integrated into clothing. This can be used to control different electronic devices.
Because these technologies create new and unusual forms of interaction, we will also look at third-party perceptions of a user’s interactions with a wearable E-textile interface. Social conventions play an important role in the development of these new forms of interaction. The willingness to make a gesture depends largely on how appropriate these interactions look and feel in public. \cite{touch-wrist} The usability of these technologies must also be examined: to support the user as efficiently, effectively and satisfactorily as possible, the designers must consider the interaction with the system in addition to the placement of these body-worn controllers.

\section{Showcase of technology examples}
%
% Maybe first a definition of body-worn controllers and a not, that we want to concentrate on Touch-Input.
% I think it is a good idea to pick a few examples which were mentioned in the papers and discuss them briefly.
% For the first part (just the showcase) it is sufficient to just write down the basic technology (with Source): So what it is? How does is work (very briefly)? What it is used for?
% We can get more into detail in the pros/cons section below
The body worn controllers\footnote{input sources made by joining small electric devices with everyday worn clothing} %footnote
allow to have bigger controlling surfaces ready to be used without needing to take out the main device. Moreover, these controllers allow users to express input gestures, which are comparable to everyday gestures. This allows the user to express emotions while using the device. If on a touchscreen, the feedback of the input is based only on vibration and visual input. With body-worn controllers, the input surface is placed over the skin, which allows direct feedback to the user, including the emotion put into performing a gesture. An example made by Wiegel et Al. %in More Than Touch
\cite[p.185]{more-touch} is “Uncomfortable interactions were chosen for actions that are very important and not reversible”, which shows how the gesture chosen to perform an action were related to the consequence that this action may lead to.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{jogwheel.png}
  \caption{\emph{Jogwheel} as presented in \protect\cite{touch-wrist}}
  \label{fig:1}
\end{figure}
%
%
% here we can add everything to e-Textiles
Another type of E-textile is the one presented by SmartSleeve \cite{smart-sleeve}, where the fabric is an input surface made of alternating conductive and insulating textile bands. Then another pressure-sensitive fabric is set under the first one, beneath which there is another fabric alternating conductive and non-conductive bands. This layer is positioned orthogonally to the top layer. This allows the fabric to register both the position and the intensity of a touch with a matrix representation. The sensor density is 1.66 sensors per square inch, which allows for a moderate sensing capability. However, since the system is working with the short-circuit principle, and the wires are not insulated, skin moisture can lead to false position sensing. To avoid this, the authors have put SmartSleeve over a long-sleeved running shirt. This solution, however, becomes unsuitable in summer for obvious reasons. %maybe better written
%
Project Jaquard \cite{jacquard}, instead, utilizes a capacitive system where the fabric is one layer textile, in which some of the wires in the weft and warp are electrical wires. As the touch sensing is made with capacitance measure, this system is not vulnerable to errors due to the skin moisture. Different materials have been investigated to create the electrical wires, both for simplicity of the creation of the fabric and the necessity of flexibility needed for comfort. The material chosen was copper because it has similar properties to cotton, silk and wool when the wire diameter is 50µm.
%
% \subsection{SimpleSkin: Towards Multipurpose Smart Garments}
The Simple Skin \cite{simpleskin} is an example of a modular smart textile garment, created for commercially viable multi-purpose garments. It combines multiple sensing modalities, namely capacitive sensing for head and wrist movement , resistive sensing for arm movement
% \cite{vogel4} #TODO
and bio-impedance sensing for heart rate. This allows optional smartphone interaction. This placement is intended to detect the angles of the joints to analyse posture, monitor vital signs, and offer input on the socially accepted region of the lower arm with the wrist sensor. Finally, the smartphone grants the options of a regular input device or a supportive device to supply the user with  information. This modular design allows for the separation of various components such as sensing, acquisition hardware, processing, and software. This allows the use of various combinations of different technologies on the same garment, and enables modifications of the underlying fabric with detachable hardware. \cite{simpleskin}

@@

The perception of technology use in public spaces can be different in every culture, because the standards of social behaviour vary in many countries. This also should be considered in the development of new interactive systems, which is why it also needs to be part of the research of social acceptance. Campbell et al. \cite{mobile-phones} have shown in their study on the public use of mobile phones, that there are indeed differences in this field. For example, the same paper just quoted, discovered that in Japan the use of mobile phones in buses or sidewalks is less accepted by society than in a restaurant. In many other cultures, this is rather the opposite. \cite{mobile-phones} These results lead to the conclusion that there are also cultural differences in perception when interacting with body-worn controllers which need to be explored.

@@

\subsection{State of social acceptability of wearables}
``...as basic needs are met, attention shifts to higher-order needs. In the case of wearable technologies, the field has matured to the point where basic needs like physical comfort, accessibility, and usability are met, higher-order needs for social acceptance and self-actualization become increasingly imperative.’’ \cite[p. 4162]{social-comfort} %(\cite{social-comfort} page 4162)
This quote underlines the current challenge Wearables have to face to succeed in the long term. Since as mentioned by Dunne, Lucy E., et al. \cite{social-comfort} the basic needs are figured out for Wearables they now have to please higher needs - such as social acceptability. To illuminate the current state some results of current researches are presented here.

@@

Result “Don’t Mind me Touching my Wrist” \cite{touch-wrist}: The "Jogwheel"  mentioned above was used in an online survey to evaluate the social acceptance of an E-textile interface by showing a demo video to the participants. The results showed gender and cultural differences in this field. There were large gender-specific differences, especially in body placement. For example, while the placement of the interface on a man's chest looked very accessible and natural to the participants, the same part of a woman's body was perceived as “disgusting”. \cite{touch-wrist} The participants from the USA found the interaction with the interface at the waist by a man more embarrassing than by a woman. The answers of the South Koreans were exactly reversed here, which indicates to cultural differences. \cite{touch-wrist}

@@

\subsection{Evaluation of placement areas for body-worn controllers }
The single most prominent factor mentioned in the respective research projects, besides the type of gesture, was the body region where the Wearable was placed. The figure \ref{fig:body} shows the preferred regions in the individual studies. When comparing the researches there are two main observations to be made. First there is a strict aversion against socially suggestive regions for Wearables. These regions include genitalia, the bottom and, especially when asking female subjects, the upper torso area, including the sternum and collarbone.
Second the all around preferred area for Wearable seems to be the forearm and wrist region - especially of the non dominant hand. \cite{touch-wrist,pinstripe,social-comfort,belt} %[1,6,7,8]
This may be due to the fact that interaction with technologies on these body regions is already established and accepted in society through other devices such as watches and fitness bracelets. However, these body regions are also easy to reach and offer sufficient space for various gestures, which makes these body regions very suitable for body-worn controllers in the future as well.
\section{Outlook on social acceptability of wearables in the future}
When looking into the future of Wearables one of the most interesting parts is the social acceptability. Since physical comfort, accessibility, and usability are for the most part figured out \cite{social-comfort}.
One, if not the most, important part of social acceptability are the performed gestures. Currently there is always a tradeoff which needs to be balanced when designing gestures, which are meant to be used in public situations. On one hand natural gestures are inherently more unobtrusive and socially acceptable. On the other hand these gestures are far more difficult to track and evaluate precisely, since they are easily confusable with everyday gestures that were not meant to be an input \cite{social-comfort}.
One approach to solve this problem is improving on better technology and tracking to avoid both confusing inputs and social discomfort at the same time. Another approach could be focussed around the social vocabulary more than around the gestures themselves.
``Because there does not currently exist a standard `vocabulary’ of gesture, it is correspondingly more difficult for viewers to match a perceived gesture with a previously-understood meaning.'' \cite[p. 4160]{social-comfort} %page 4160
Looking at this quote the other solution seems obvious. With the further adaptation of a novel technology there comes a better understanding of the gestures and finally a new social vocabulary regarding it. There were similar developments when the bluetooth headset \cite{social-comfort} or the Sony Walkman \cite{touch-wrist} were established.
In conclusion, the social acceptance of body-worn controllers can be increased through further developments and by bringing them more and more in contact with the public.
